---
title: "Analysis scripts: K. pneumoniae in QECH"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Analysis scripts: K. pneumoniae in QECH}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```



```{r setup}
library(QECHospitalKlebs)
```





## Introduction

This document generates the tables and figures for the manuscript:

<br />

*Longitudinal analysis within one hospital in sub-Saharan Africa over 20 years reveals repeated replacements of dominant clones of _Klebsiella_ _pneumoniae_ and stresses the importance to include temporal patterns for vaccine design considerations*

<br />

Eva Heinz^1,2^, Oliver Pearse^2,3^, Allan Zuza^3^, Sithembile Bilima^3^, Chisomo Msefula^4^, Patrick Musicha^2,3^, Patriciah Siyabu^5^, Edith Tewesa^5^, Fabrice E Graf^2^, Rebecca Lester^3,6^, Samantha Lissauer^3,7^, Jennifer Cornick^3,7^, Joseph M Lewis^2,3,7^, Kondwani Kawaza^3,4^, Nicholas R Thomson^8,9^, Nicholas A Feasey^2,3,10^

1. Department of Vector Biology, Liverpool School of Tropical Medicine, Liverpool, UK
2. Department of Clinical Sciences, Liverpool School of Tropical Medicine, Liverpool, UK
3. Malawi Liverpool Wellcome Programme, Kamuzu University of Health Sciences, Blantyre, Malawi
4. Kamuzu University of Health Sciences, Blantyre, Malawi
5. Queen Elizabeth Central Hospital, Blantyre, Malawi
6. Division of Infection & Immunity, University College London, UK
7. Department of Clinical Infection, Microbiology and Immunology, University of Liverpool, Liverpool, UK
8. Parasites and Microbes Program, Wellcome Sanger Institute, Hinxton, UK
9. London School of Hygiene and Tropical Medicine, London, UK
10. School of Medicine, St Andrews University, UK


## STs over time, wards over time

```{r metadata over time}


```



## Antimicrobial resistance plots

```{r AMR plots}


```



## Antimicrobial resistance and virulence over time

```{r AMR VIR over time}


```



## Full ariba predictions mapped to trees

```{r ariba trees}


```




## Clustering of closely related isolates, tree visualisations

```{r rPineCone cluster}


```




## O- and K-type figures

```{r O-K-types}


```





### Reproducability, session info

```{r sessinfo}

sessionInfo()

```

